Discrete Quantum Theories and Computing

### 2019-02-22 dissertation.tex
dissertation.lyx Sec. 4.7 Reviewed...

Although cardinal probabilities can be ordered as discussed in \ref{discretequantumtheoryIIb},
it is still unclear how to define their arithmetic operations, which
is required to further define the mixed states and expectation values.
Notice that the cardinal probability over finite fields defined in
Eq.~(\ref{eq:cardinalProbability}) is based on the conventional
Born rule, Eq.~(\ref{eq:Born-ComputationalBasis}) which could be
derived axiomatically according to Gleason's theorem as in Sec.~\ref{subsec:Gleason's-Theorem-and}.
\begin{comment}
Although they are valid inside an ordered region, some quantum algorithms,
like Shor's~\cite{Shor1999,Mermin2007,Jaeger2007}, don't localize
their computation in a small region. \todo{Remove the discussion
of Shor's algorithm if I cannot verify it on DQT~(I) on time. }Moreover, 
\end{comment}
To overcome these issues, we want to follow the steps in Sec.~\ref{sec:fuzzy}
to define events and probability measures, and see whether we could
have a Gleason-like theorem which induces a correspondence between
states and probability measures qualified as a discrete Born rule.

@@

After defining a QPMFF~$\mu$, we would like to follow the idea of
Gleason's theorem to see whether we might trace back from $\mu$ to
a pure or mixed state and summarize this correspondence as a discrete
Born rule. When $p=D=3$, there is indeed some correspondence between
QPMFFs and states. However, when $D=3$ and $p=7$, we have numerically
verified that the only QPMFF $\mu\colon\events_{7^{2}}\rightarrow\left[0,1\right]$
is %
\begin{comment}
\todo{Equal probable.}
\end{comment}
\begin{equation}
\mu\left(P\right)=\begin{cases}
0 & \textrm{if }P=\mathbb{0}\,;\\
\frac{1}{3} & \textrm{if }P\textrm{ is a one-dimensional projector;}\\
\frac{2}{3} & \textrm{if }P\textrm{ is a two-dimensional projector;}\\
1 & \textrm{if }P=\mathbb{1}\,.
\end{cases}\label{eq:QPMFF37}
\end{equation}
As we computed in Eq.~(\ref{eq:numberOfIrreducibleStates}), the
number of unique irreducible pure states in $\ff{7^{2}}^{3}$ is 
\begin{equation}
\frac{7^{3-1}\left(7^{3}-\left(-1\right)^{3}\right)}{7+1}=2107\label{eq:numberOfIrreducibleStates-7-3}
\end{equation}
which is much more than one, the number of QPMFF\@. Since we don't
have enough QPMFF to correspond even pure states, there is no discrete
Born rule in this case. In general, we conjecture that QPMFF is always
unique for $D\ge3$ except $D=p=3$.

@@

Following the similar computation to the previous paragraph, we now
compute the probabilities induced by $\ket{\Phi'}=U'\ket{\oplus}=\frac{2\alpha}{\gamma}\ket{0}+\frac{1}{\gamma}\ket{2}$,
where $\alpha=\fnorm*{\frac{1}{2}}\in\ff{p^{2}}$ and
\begin{equation}
U'=\begin{pmatrix}\alpha & \alpha & 0\\
-\alpha & \alpha & 0\\
0 & 0 & 1
\end{pmatrix}\,.
\end{equation}
Since $\ket{2}$ is invariant under $U'$, Eq.~(\ref{eq:basis-independent})
implies $\muF_{\oplus}\left(P_{2}\right)=\muF_{U'\ket{\oplus}}\left(U'P_{2}U'^{\dagger}\right)=\muF_{\Phi'}\left(P_{2}\right)$.
Also, Eq.~(\ref{eq:zero}) implies $\muF_{\Phi'}\left(P_{0}\right)+\muF_{\Phi'}\left(P_{2}\right)=\muF_{\Phi'}\left(P_{0}+P_{2}\right)=1$.
By moving $\muF_{\Phi'}\left(P_{2}\right)$ to the right-hand side
of the equation, we have 
\begin{equation}
\muF_{\Phi'}\left(P_{0}\right)=1-\muF_{\Phi'}\left(P_{2}\right)=1-\muF_{\oplus}\left(P_{2}\right)=1-\tfrac{1}{3}=\tfrac{2}{3}\,.\label{eq:prob23}
\end{equation}

>>>>

Although cardinal probabilities can be ordered as discussed in Sec.~\ref{discretequantumtheoryIIb},
it is still unclear how to define their arithmetic operations, which
is required to further define the mixed states and expectation values.
Notice that the cardinal probability over finite fields defined in
Eq.~(\ref{eq:cardinalProbability}) is based on the conventional
Born rule, Eq.~(\ref{eq:Born-ComputationalBasis}), which could be
derived axiomatically according to Gleason's theorem as in Sec.~\ref{subsec:Gleason's-Theorem-and}.
\begin{comment}
Although they are valid inside an ordered region, some quantum algorithms,
like Shor's~\cite{Shor1999,Mermin2007,Jaeger2007}, don't localize
their computation in a small region. \todo{Remove the discussion
of Shor's algorithm if I cannot verify it on DQT~(I) on time. }Moreover, 
\end{comment}
To overcome these issues, we want to follow the steps in Sec.~\ref{sec:fuzzy}
to define events and probability measures, and see whether we could
have a Gleason-like theorem which induces a correspondence between
states and probability measures qualified as a discrete Born rule.

@@

After defining a QPMFF~$\mu$, we would like to follow the idea of
Gleason's theorem to see whether we might trace back from $\mu$ to
a pure or mixed state and summarize this correspondence as a discrete
Born rule. When $p=D=3$, there is indeed some correspondence between
QPMFFs and states. However, when $D=3$ and $p=7$, we have numerically
verified that the unique QPMFF $\mu\colon\events_{7^{2}}\rightarrow\left[0,1\right]$
is the equal probable one
\begin{equation}
\mu\left(P\right)=\begin{cases}
0 & \textrm{if }P=\mathbb{0}\,;\\
\frac{1}{3} & \textrm{if }P\textrm{ is a one-dimensional projector;}\\
\frac{2}{3} & \textrm{if }P\textrm{ is a two-dimensional projector;}\\
1 & \textrm{if }P=\mathbb{1}\,.
\end{cases}\label{eq:QPMFF37}
\end{equation}
However, the irreducible pure states in $\ff{7^{2}}^{3}$ is not unique.
Indeed, there are 
\begin{equation}
\frac{7^{3-1}\left(7^{3}-\left(-1\right)^{3}\right)}{7+1}=2107\label{eq:numberOfIrreducibleStates-7-3}
\end{equation}
irreducible pure states in $\ff{7^{2}}^{3}$ as we computed in Eq.~(\ref{eq:numberOfIrreducibleStates}).
Since we don't have enough QPMFF to correspond even pure states, there
is no discrete Born rule in this case. In general, we conjecture that
QPMFF is always unique for $D\ge3$ except $D=p=3$.

@@

Following the computation similar to the previous paragraph, we now
compute the probabilities induced by $\ket{\Phi'}=U'\ket{\oplus}=\frac{2\alpha}{\gamma}\ket{0}+\frac{1}{\gamma}\ket{2}$,
where $\alpha=\fnorm*{\frac{1}{2}}\in\ff{p^{2}}$ and
\begin{equation}
U'=\begin{pmatrix}\alpha & \alpha & 0\\
-\alpha & \alpha & 0\\
0 & 0 & 1
\end{pmatrix}\,.
\end{equation}
Since $\ket{2}$ is invariant under $U'$, Eq.~(\ref{eq:basis-independent})
implies $\muF_{\oplus}\left(P_{2}\right)=\muF_{U'\ket{\oplus}}\left(U'P_{2}U'^{\dagger}\right)=\muF_{\Phi'}\left(P_{2}\right)$.
Also, Eq.~(\ref{eq:zero}) implies $\muF_{\Phi'}\left(P_{0}\right)+\muF_{\Phi'}\left(P_{2}\right)=\muF_{\Phi'}\left(P_{0}+P_{2}\right)=1$.
By moving $\muF_{\Phi'}\left(P_{2}\right)$ to the right-hand side
of the equation, we have 
\begin{equation}
\muF_{\Phi'}\left(P_{0}\right)=1-\muF_{\Phi'}\left(P_{2}\right)=1-\muF_{\oplus}\left(P_{2}\right)=1-\tfrac{1}{3}=\tfrac{2}{3}\,.\label{eq:prob23}
\end{equation}
